### 2025-01-21 paper/icml25/first/sects/shacpp.tex
refactoring

As mentioned previously, we aim to address the limitations of SHAC by approximating the gradients of a differentiable environment ($R$ and $F$) with neural networks ($R_\theta$ and $F_\theta$). In particular, we aim to train a neural network to approximate the reward function by minimizing: 
\begin{align*}
    \mathbb{E}_{\tau\sim\pi}\left[\left\vert\left\vert \frac{1}{\horizon\na}\sum_{j=0}^{\horizon-1} \sum_{i\in\agents} {\rewards{i}}_\theta(s_j^\tau, a_j^\tau, s_{j+1}^\tau) \right.\right.\right. -\\ \left.\left.\left. \rewards{i}(s_j^\tau, a_j^\tau, s_{j+1}^\tau) \right\vert\right\vert^2\right] 
\end{align*}

In practice, this is achieved by leveraging a \emph{replay buffer} to cache the state-action-reward triplets obtained from the environment. The neural network is then trained to minimize the mean squared error (MSE) loss. Specifically, we minimize the loss $\mathcal{L}_\theta^{\rewards{i}}:(\states\times\actions{i}\times\states)^N\rightarrow\mathbb{R}$:
$$ \mathcal{L}_\theta^{\rewards{i}}(\batch) = \frac{1}{\bs}\sum_{s,a,r \in \batch}\left|\left| {\rewards{i}}_\theta(s, a) - r \right|\right|^2 $$
Where $B$ represents a batch of state-action-reward triplets sampled uniformly from the cache.

Similarly, we aim to train a neural network to approximate the transition function. However, to avoid vanishing or exploding gradients, we employ a non-recursive approach similar to Action World Model~\cite{Ma24}. This corresponds to train a neural network that given a starting observation from $\mathcal{S}$ and a sequence of $\mh$ actions from $(\{\actions{i}\}_{i\in\agents})^\mh$, predicts the next $\mh$ observations from $\states^\mh$. We aim to minimize:
%\sum_{j=0}^{\scriptscriptstyle\mh-1}
\begin{align*}
    \mathbb{E}_{\tau\sim\pi}\left[\frac{1}{\bs\mh}\sum_{t=0}^{\scriptscriptstyle\horizon-\mh-1}\left|\left| \transition_{\theta}(s_t^\tau, \{a^\tau_t\}_{i}^{\na},\dots,\{a^\tau_{\scriptscriptstyle t+H-1}\}_{i}^{\na}) \right.\right.\right.-\\\left.\left.\left. \left[s_t^\tau,\dots,s_{t+h}^\tau\right] \right|\right|^2\right] 
\end{align*}

In practice, this is achieved by caching trajectories and then training the neural network to minimize the MSE loss. That is, we minimize the loss $\mathcal{L}_\theta^{F}:\mathcal{T}^N\rightarrow\mathbb{R}$:
$$ \frac{1}{\bs\mh}\sum_{\tau \in B}\left|\left| \transition_{\theta}(s_0^\tau, \{a_1^\tau\}_i^\na, \dots, \{a_{\mh}^\tau\}_i^\na) - \left[ s^\tau_1, \dots, s^\tau_\mh \right] \right|\right|^2 $$
Where $B$ represents a batch of trajectories of length $\mh$ sampled uniformly from the cache.

Further, we modify the SHAC policy loss, $\mathcal{L}_\theta^{\pi}$, to include a term that penalizes actions outside the action space. Suppose the action space is $(l,r)$ (with $l<r$) for all agents, we minimize the loss $\mathcal{L'}_\theta^\pi:$
\begin{figure*}[!t]
    \centering
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{figs/dispersion.png}
        \caption{Dispersion}
        \label{fig:dispersion}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{figs/transport.png}
        \caption{Transport}
        \label{fig:transport}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{figs/discovery.png}
        \caption{Discovery}
        \label{fig:discovery}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{figs/sampling.png}
        \caption{Sampling}
        \label{fig:sampling}
    \end{subfigure}
    \caption{VMAS environments used in the experiments.}\label{fig:scenarios}
\end{figure*}

$$\mathcal{L}_\theta^\pi(\batch) + \alpha \frac{1}{\bs\horizon}\sum_{\tau\in\batch}\sum_{a\in\tau} d(a^\tau,(a,b))$$
$$d(x,(l,r)) = \begin{cases}(l-x)^2 & \text{if } x < l \\ (x-r)^2 & \text{if } x > e \\ 0 & \text{otherwise} \end{cases}$$

>>>>

As mentioned previously, we aim to address the limitations of SHAC by approximating the gradients of a differentiable environment ($R$ and $F$) with neural networks ($R_\theta$ and $F_\theta$). In particular, we aim to train a neural network to approximate the reward function by minimizing: 
\begin{align*}
    \underset{\tau\sim\pi}{E}\left[\hspace{-.7cm}\sum_{\hspace{.7cm}\scriptscriptstyle s_i,a_{ij},s_{i+1}\in\tau} \hspace{-.8cm}\left\vert\left\vert {\rewards{i}}_\theta(s_i, a_{ij}, s_{i+1}) -\rewards{i}(s_i, a_{ij}, s_{i+1}) \right\vert\right\vert^2\right] 
\end{align*}

In practice, this is achieved by leveraging a \emph{replay buffer} to cache the state-action-reward triplets obtained from the environment. The neural network is then trained to minimize the mean squared error (MSE) loss. Specifically, we minimize the loss $\mathcal{L}_\theta^{\rewards{i}}:(\states\times\actions{i}\times\states)^N\rightarrow\mathbb{R}$:
$$ \mathcal{L}_\theta^{\rewards{i}}(\batch) = \frac{1}{\bs}\sum_{s,a,s',r \in \batch}\left|\left| {\rewards{i}}_\theta(s, a) - r \right|\right|^2 $$
Where $B$ represents a batch of state-action-state-reward tuples sampled uniformly from the cache.

Similarly, we aim to train a neural network to approximate the transition function. However, to avoid vanishing or exploding gradients, we employ a non-recursive approach similar to Action World Model~\cite{Ma24}. This corresponds to train a neural network that given a starting observation from $\mathcal{S}$ and a sequence of $\mh$ actions from $\alla^\mh$, predicts the next $\mh$ observations from $\states^\mh$. We aim to minimize:

\begin{align*}
    \underset{\tau\sim\pi}{E}\left[\hspace{-1.2cm}\sum_{\scriptscriptstyle \hspace{1cm}\substack{s_t,\dots,s_{t+h+1},\\\hspace{.3cm}a_{t*},\dots,a_{t+h*}\in\tau}}\hspace{-1.2cm} \left|\left| \transition_{\theta}(s_t, a_{t*},\dots,a_{\scriptscriptstyle i+h*}) - [s_t,\dots,s_{\scriptscriptstyle t+h+1}] \right|\right|^2\right] 
\end{align*}

Where $a_{t*}=\{a_{ti}\}_{i\in\agents}$. In practice, this is achieved by caching trajectories and then training the neural network to minimize the MSE loss. That is, we minimize the loss $\mathcal{L}_\theta^{F}:\mathcal{T}^N\rightarrow\mathbb{R}$:
$$ \frac{1}{\bs\mh}\sum_{\tau \in B}\left|\left| \transition_{\theta}(s_0^\tau, a_{1*}^\tau, \dots, a_{\mh*}) - \left[ s^\tau_1, \dots, s^\tau_\mh \right] \right|\right|^2 $$
Where $B$ represents a batch of trajectories of length $\mh$ sampled uniformly from the cache.

Further, we modify the SHAC policy loss, $\mathcal{L}_\theta^{\pi}$, to include a term that penalizes actions outside the action space. Suppose the action space is $(l,r)$ (with $l<r$) for all agents, we minimize the loss $\mathcal{L'}_\theta^\pi:$

$$\mathcal{L}_\theta^\pi(\batch) + \alpha \frac{1}{\bs\horizon}\sum_{\tau\in\batch}\sum_{a\in\tau} d(a^\tau,(a,b))$$
$$d(x,(l,r)) = \begin{cases}(l-x)^2 & \text{if } x < l \\ (x-r)^2 & \text{if } x > r \\ 0 & \text{otherwise} \end{cases}$$
